---
title: "HW3 - Intro to Tidyverse"
author: Kayla Kippes
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Task 1

### Part A

We cannot use the read_csv() to read in these data because the delimiter is a semicolon and not a comma.

```{r}
library(tidyverse)
library(palmerpenguins)

data <- read_csv2("data/data.txt")
data
```

### Part B

```{r}
data2 <- read_delim("data/data2.txt", delim = "6", col_types = "fdc")
data2
```

## Task 2

### Part A

```{r}
trailblazer <- read_csv("data/trailblazer.csv")
trailblazer
```

### Part B

```{r}
library(tidyr)
#Put the data into long format and tidy up columns
trailblazer_longer <- trailblazer |>
  pivot_longer(cols = 2:11, 
               names_to = c("Game", "Location"),
               names_sep = "_",
               values_to = "Points")
trailblazer_longer[1:5,]
```

### Part C

```{r}
trailblazer_longer |>
  pivot_wider(names_from = Location, values_from = Points, values_fill = 0) |>
  group_by(Player) |>
  summarise(Mean_Home = mean(Home, na.rm = TRUE),
            Mean_Away = mean(Away, na.rm = TRUE),
            Difference = Mean_Home - Mean_Away) |>
  arrange(desc(Difference))
```

Jusuf Nurkic, Robert Covington, Nassir Little, Damian Lillard, and Cody Zeller all scored more points at home than away in the first 10 games.

## Task 3

### Part A

<NULL> means that there are no values for that species and island combination.

<dbl [52]> means that the list contains 52 values in it.

<list> means that the output contains a list-column.

### Part B

```{r}
penguins |>
   select(species, island, bill_length_mm) |>
  group_by(species, island) |>
  summarise(count = n()) |>
  pivot_wider(names_from = island, values_from = count, values_fill = 0) 
```

## Task 4

```{r}
penguins |>
  select(species, bill_length_mm) |>
  mutate(bill_length_mm = case_when(
    is.na(bill_length_mm) & species == "Adelie" ~ 26,
    is.na(bill_length_mm) & species == "Gentoo" ~ 30,
    TRUE ~ bill_length_mm)) |>
  arrange(bill_length_mm) |>
  print(n = 10)
```

